Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-07 csf-2017/paper.tex
lr editing

\subsubsection{Future worlds}
In order to model changes to memory over time, we use the future world
relations. The future world relations model how more memory may be
allocated but also how memory may change. How memory is allowed to
change over time depends on the region that governs it and the future
world relation used. The most interesting part about the future worlds
is how it affects the regions, so we first look at the properties of
the \emph{public future region} relation and the \emph{private future
  region} relation. We then use these relations to describe the future
world relations.

{\bf Private future regions}
In the private future region relation, we allow revocation of $\temp$
regions. The regions cannot be revoked by removing them from the
world, because we need the future world relations to be
monotone. This is why we have the $\revoked$ region, so we have
a region that can serve as a mask. We can, however, let any region
serve as a mask, so in the private future region relation, we allow a
$\temp$ region to become any valid region. As the $\revoked$ region
serves as a mask, we let any valid region take its place in the
private future region relation.

If a $\temp$ region is not masked by a region, then it behaved like a
$\perm$ region. The $\perma$ regions stay the same type and they are
allowed to take a transition in the private part of the transition
system. Which can be thought of as the memory changes according to
some protocol.

The properties of the \emph{private future region} relation are:
\begin{mathpar}
  \inferrule{  (s,s') \in \phi \\
    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}

{\bf Public future regions}
In the public future regions we do not permit revocation of $\temp$
regions. In this relation, the $\temp$ and $\perma$ regions must stay
the same type and they are allowed to take a transition in the public
part of the transition system.

The $\revoked$ regions serve as masking regions, but are not as
liberal as in the private future region relation when it comes to
letting other regions replace them as masks. In the public future
region relation, we only allow $\revoked$ and $\temp$ regions to take
on the job as masking region. We have this limitation, because we
want the private future region relation to be able to reuse any region
it has revoked.

The \emph{public future} region relations satisfies the following properties:
\begin{mathpar}
  \inferrule{  (s,s') \in \phi_\pub \\
    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}

Notice that the public future region relation is a subset of the
private future region relation, so anything you are allowed to do in
the public future region relation, the private future region relation
can do the same.

% The future world relation are really the ones from thm, the above
% are properties of them.
{\bf Future world relations}
There are two future world relations a \emph{public future world}
relation and a \emph{private future world} relation. In both future
world relations, we have extension ordering, so a future world $W'$ of
$W$ has at least the same regions as $W$, but $W'$ may have new regions
imposing new protocols on parts of the memory. The two future world
relations differ in the way existing regions are allowed to
evolve. The public future world relation uses the public future region
relation and the private future world relation uses the private future
region relation:
\begin{mathpar}
  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
  { W' \futurewk W }
  \and
  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
  { W' \futurestr W }
\end{mathpar}

% Relate to STSs for high-level languages. Have the public-private
% transitions, but they are also used for more than this
\todo[inline]{Relate to STSs for high-level languages?}

>>>>

\subsubsection{Future worlds}
The future world relations model how memory may evolve over time. 
The \emph{public future world} 
$W' \futurewk W$ is defined to mean that $\dom(W') \supseteq \dom(W)$
and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$.  That is,
in a public future world, more regions may have been allocated, and
each of the regions may also have evolved according to the public future
region relation (defined below). The \emph{private future world} relation
$W' \futurestr W$ is defined similarly, using a private future region
relation.

The properties of the \emph{private future region} relation are:
\begin{mathpar}
  \inferrule{  (s,s') \in \phi \\
    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
These rules express that, in the private future region relation, 
we allow revocation of temporary regions. 
As explained above, this can be done using a special revoked region,
which does not desribe any memory, but we also allow  
any other region to serve as a mask, and thus any region $r$ can be a
future region of a temporary region and of a revoked region.
If a $\temp$ region is not masked by a region, then it behaves like a
$\perm$ region. The $\perma$ regions stay permanent and they are
allowed to take a transition in the private part of the transition
system.

The \emph{public future} region relations satisfies the following properties:
\begin{mathpar}
  \inferrule{  (s,s') \in \phi_\pub \\
    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
In the public future regions we do not permit revocation of temporary
regions. Instead, temporary and permanent regions are only 
allowed to take a transition in the public part of the transition system.

The $\revoked$ regions serve as masking regions, but are not as
liberal as in the private future region relation when it comes to
letting other regions replace them as masks. In the public future
region relation, we only allow $\revoked$ and $\temp$ regions to take
on the job as masking region. We have this limitation, because we
want the private future region relation to be able to reuse any region
it has revoked.
\todo[inline]{can we make this crisper ?}

Notice that the public future region relation is a subset of the
private future region relation.

%%%%%%%%

% In order to model changes to memory over time, we use the future world
% relations. The future world relations model how more memory may be
% allocated but also how memory may change. How memory is allowed to
% change over time depends on the region that governs it and the future
% world relation used. The most interesting part about the future worlds
% is how it affects the regions, so we first look at the properties of
% the \emph{public future region} relation and the \emph{private future
%   region} relation. We then use these relations to describe the future
% world relations.

% {\bf Private future regions}
% In the private future region relation, we allow revocation of $\temp$
% regions. The regions cannot be revoked by removing them from the
% world, because we need the future world relations to be
% monotone. This is why we have the $\revoked$ region, so we have
% a region that can serve as a mask. We can, however, let any region
% serve as a mask, so in the private future region relation, we allow a
% $\temp$ region to become any valid region. As the $\revoked$ region
% serves as a mask, we let any valid region take its place in the
% private future region relation.

% If a $\temp$ region is not masked by a region, then it behaved like a
% $\perm$ region. The $\perma$ regions stay the same type and they are
% allowed to take a transition in the private part of the transition
% system. Which can be thought of as the memory changes according to
% some protocol.

% The properties of the \emph{private future region} relation are:
% \begin{mathpar}
%   \inferrule{  (s,s') \in \phi \\
%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
%   {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ r \in \Regions }
%   { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ r \in \Regions }
%   { r \futurestr \revoked }
% \end{mathpar}

% {\bf Public future regions}
% In the public future regions we do not permit revocation of $\temp$
% regions. In this relation, the $\temp$ and $\perma$ regions must stay
% the same type and they are allowed to take a transition in the public
% part of the transition system.

% The $\revoked$ regions serve as masking regions, but are not as
% liberal as in the private future region relation when it comes to
% letting other regions replace them as masks. In the public future
% region relation, we only allow $\revoked$ and $\temp$ regions to take
% on the job as masking region. We have this limitation, because we
% want the private future region relation to be able to reuse any region
% it has revoked.

% The \emph{public future} region relations satisfies the following properties:
% \begin{mathpar}
%   \inferrule{  (s,s') \in \phi_\pub \\
%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
%   {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
%   { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
%   \and
%   \inferrule{ }
%   { \revoked \futurewk \revoked }
% \end{mathpar}

% Notice that the public future region relation is a subset of the
% private future region relation, so anything you are allowed to do in
% the public future region relation, the private future region relation
% can do the same.

% % The future world relation are really the ones from thm, the above
% % are properties of them.
% {\bf Future world relations}
% There are two future world relations a \emph{public future world}
% relation and a \emph{private future world} relation. In both future
% world relations, we have extension ordering, so a future world $W'$ of
% $W$ has at least the same regions as $W$, but $W'$ may have new regions
% imposing new protocols on parts of the memory. The two future world
% relations differ in the way existing regions are allowed to
% evolve. The public future world relation uses the public future region
% relation and the private future world relation uses the private future
% region relation:
% \begin{mathpar}
%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
%     \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
%   { W' \futurewk W }
%   \and
%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
%     \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
%   { W' \futurestr W }
% \end{mathpar}

% % Relate to STSs for high-level languages. Have the public-private
% % transitions, but they are also used for more than this
% \todo[inline]{Relate to STSs for high-level languages?}
